Supplementary Material

### 2024-07-31 manuscript/manuscript.qmd
add pcls data

Building on our *in vitro* findings indicating an antifibrotic effect of lactate transporter inhibition, we proceeded to evaluate the efficacy of MCT inhibitors in a bleomycin-induced mouse model of pulmonary fibrosis. Mice received 1.2 U/kg bleomycin by intratracheal administration. Seven days later, the animals began treatment with AZD3965 (100 mg/kg twice daily) or VB124 (30 mg/kg once daily) or vehicle by oral gavage (**Fig. 6A**). Considering likely toxicity, we did not assess the combination of inhibitors. Compared to vehicle, mice treated with VB124 had increased weight recovery 21 days after bleomycin administration (**Fig. S8**). Lung mechanics improved by approximately 50% of baseline compared to vehicle controls following 14 days of MCT inhibitor treatment (**Fig. 6B-C**). This improvement in measures of lung stiffness was further supported by histologic assessment of fibrosis severity using Ashcroft scoring [@ashcroft1988] (**Fig. 6D-E**) and hydroxyproline content measurement (**Fig. 6F**). Taken together, these data demonstrate substantially decreased pulmonary fibrosis severity after 14 days of treatment with lactate transport inhibitors.

@@

![**Fig. 6. Lactate transporter inhibition attenuates bleomycin-induced pulmonary fibrosis.** (**A**) C57Bl/6N mice were administered bleomycin (1.2 U/kg) on day 0. Beginning on day 7, animals were treated with AZD3965 (AZD) or VB124 (VB) for 14 days prior to euthanasia on day 21. On day 21, mice were anesthetized with pentobarbital for lung function measurements. (**B**) Static compliance of the lung was measured by pressure-volume loop analysis. (**C**) Elastance was measured using the forced oscillation technique for 8 s (Prime-8) (N = 13-17). (**D**) Representative trichrome-stained histologic sections of mouse lungs. Scale bar is 200 μm. (**E**) Ashcroft scores from histologic sections from (D) determined by a lung pathologist blinded to treatment assignment (N = 6-7). (**F**) Right lung hydroxyproline content (N = 10-13). Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/F_06.png)

>>>>

Building on our *in vitro* findings indicating an antifibrotic effect of lactate transporter inhibition, we proceeded to evaluate the efficacy of MCT inhibitors in a bleomycin-induced mouse model of pulmonary fibrosis. Mice received 1.2 U/kg bleomycin by intratracheal administration. Seven days later, the animals began treatment with AZD3965 (100 mg/kg twice daily) or VB124 (30 mg/kg once daily) or vehicle by oral gavage (**Fig. 6A**). Considering likely toxicity, we did not assess the combination of inhibitors. Compared to vehicle, mice treated with VB124 had increased weight recovery 21 days after bleomycin administration (**Fig. S8**). Lung mechanics improved by approximately 50% of baseline compared to vehicle controls following 14 days of MCT inhibitor treatment (**Fig. 6B-C**). This improvement in measures of lung stiffness was further supported by histologic assessment of fibrosis severity using Ashcroft scoring [@ashcroft1988] (**Fig. 6D-E**) and hydroxyproline content measurement (**Fig. 6F**). Taken together, these data demonstrate substantially decreased pulmonary fibrosis severity after 14 days of treatment with lactate transport inhibitors. Similarly, dual MCT inhibition decreased TGFβ-stimulated extracellular matrix protein production in human precision-cut lung slices (**Fig. 6G**).

@@

![**Fig. 6. Lactate transporter inhibition attenuates experimental pulmonary fibrosis.** (**A**) C57Bl/6N mice were administered bleomycin (1.2 U/kg) on day 0. Beginning on day 7, animals were treated with AZD3965 (AZD) or VB124 (VB) for 14 days prior to euthanasia on day 21. On day 21, mice were anesthetized with pentobarbital for lung function measurements. (**B**) Static compliance of the lung was measured by pressure-volume loop analysis. (**C**) Elastance was measured using the forced oscillation technique for 8 s (Prime-8) (N = 13-17). (**D**) Representative trichrome-stained histologic sections of mouse lungs. Scale bar is 200 μm. (**E**) Ashcroft scores from histologic sections from (D) determined by a lung pathologist blinded to treatment assignment (N = 6-7). (**F**) Right lung hydroxyproline content (N = 10-13). (**G**) Human precision-cut lung slices were stimulated with TGFβ (5 ng/mL) and treated with AZD3965 and VB124. After 6 d of treatment, protein lysates were analyzed for extracellular matrix proteins by immunoblot. Dual MCT inhibition decreased fibronectin (FN1) and Col1a2 production. Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value < 0.05 compared to bleomycin-treated (Bleo) or TGFβ-treated vehicle controls.](figs/F_06.png)

### 2024-07-31 manuscript/supplement.qmd
add pcls data

>>>>

## Precision-cut lung slices

Human precision-cut lung slices (PCLS) were purchased from Anabios (240326LuHA, 26 y.o. female without lung disease). The slices were incubated for 48 h in DMEM/F12 with 10% serum and then treated with TGFβ (5 ng/mL) and AZD3965 (100 nM) plus VB124 (10 μM) for 6 d, exchanging the medium every 48 h with fresh TGFβ and inhibitors. PCLS were lysed in RIPA buffer containing protease inhibitors, homogenized by bead-beating, and analyzed for extracellular matrix production by immunoblot.
